---
title: "Permutation Test on Box Office Success Between Genres"
author: "Natalie Roston"
format: html
---

**Introduction**

In this analysis, I wanted to test whether there is a significant difference in the average box office revenue between action and comedy movies using the \`movies\` dataset from OpenIntro. I use a permutation test to evaluate if any observed difference in average box office gross is statistically significant under the null hypothesis that genre does not affect revenue. I hypothesize that action movies have a higher box office revenue than comedies.

**Data Setup**

```{r, message=FALSE, warning=FALSE}
library(openintro)
library(ggplot2)
library(purrr)
library(dplyr)

data("movies")

str(movies)
```
The above code sets up the data and allows me to understand its structure. 

```{r}
movies_clean <- movies |>
  filter(!is.na(box_office)) 
```

```{r}
observed_diff <- movies_clean |>
  group_by(genre) |>
  summarise(mean_box_office = mean(box_office, na.rm = TRUE)) |>
  filter(genre %in% c("action", "comedy")) |>
  summarize(observed_diff = 
    mean_box_office[genre == "action"] - mean_box_office[genre == "comedy"]
  )

observed_diff

```

```{r}
permute_diff <- function(data) {
  data |>
    mutate(shuffled_genre = sample(genre)) |>
    summarize(
      observed_diff = mean(box_office[shuffled_genre == "action"], na.rm = TRUE) - 
                      mean(box_office[shuffled_genre == "comedy"], na.rm = TRUE)
    ) |>
    pull(observed_diff) 
}
```

```{r}
observed_diff <- as.numeric(observed_diff)

set.seed(4747) 
num_simulations <- 10000
perm_diffs <- map_dbl(1:num_simulations, ~ permute_diff(movies))
p_value <- mean(abs(perm_diffs) >= abs(observed_diff))
p_value
```

```{r}
perm_data <- data.frame(perm_diffs = perm_diffs)

ggplot(perm_data, aes(x = perm_diffs)) +
  geom_histogram(bins = 30, color = "black", fill = "skyblue", alpha = 0.7) + geom_vline(xintercept = observed_diff, color = "red", linetype = "dashed", linewidth = 1) +
  labs(
    title = "Permutation Test: Box Office Gross Differences Between Action and Comedy Movies",
    x = "Difference in Mean Gross (Action - Comedy)",
    y = "Frequency"
  ) +
  theme_minimal()

```

This plot illustrates the results of a permutation test designed to assess whether there is a statistically significant difference in box office revenue between action and comedy movies. In the histogram, we see the distribution of mean differences in box office revenue under the null hypothesis, which assumes there is no true effect of genre on box office performance. Each bar represents the frequency of mean differences obtained by randomly shuffling genre labels (action and comedy) across all movies, then calculating the difference in means for each shuffle. This simulated distribution of differences reveals the range of values we might expect if genre had no impact on revenue.

The red dashed line in the plot marks the observed difference in box office revenue between action and comedy movies as calculated from the original data set. This observed difference provides the basis for testing our hypothesis. If this red line lies well outside the central bulk of the histogram, it indicates that the observed difference is unusual compared to what we would expect under the null hypothesis. In other words, a position far from the central peak of the histogram suggests that the observed difference in revenue is not easily explainable by random variation alone.

```{r}
p_value <- mean(abs(perm_diffs) >= abs(observed_diff))
p_value

```

The position of the observed difference near the tail of the distribution aligns with a p-value of 0.0386, indicating a statistically significant result. This means that, under the null hypothesis, there is only a 3.86% chance of observing a difference as extreme as the one in the data set. Consequently, the plot supports the conclusion that there is likely a true difference in box office revenue between action and comedy movies, with genre possibly influencing financial performance. This visualization, therefore, provides both a statistical and a visual basis for concluding that genre may have an impact on box office success.

The permutation test and histogram assess whether the observed difference in box office revenue between action and comedy movies (31 million dollars) is statistically significant or likely due to random chance. The histogram shows the null distribution, representing what differences we might expect if genre had no impact. With a p-value of 0.0386, the observed difference is unlikely to have occurred by chance, suggesting a real difference in box office revenue between action and comedy movies in the broader population. This result supports the idea that action movies tend to gross more than comedies on average.

Source:

OpenIntro. (2021). movies dataset. In OpenIntro: Data sets for introductory statistics (Version 3.4). Retrieved from https://www.openintro.org
