---
title: "Project 2"
description: |
  Text Analysis
author: Natalie Roston
date: September 18, 2024
format: html
execute:
  warning: false
  message: false
---

```{r}
library(tidytuesdayR)
library(dplyr)
library(stringr)
library(ggplot2)

tuesdata <- tidytuesdayR::tt_load('2021-04-20')
netflix <- tuesdata$netflix

netflix |>
  group_by(type, rating) |>
  summarize(count = n()) |>
  ggplot(aes(x = rating, y = count)) + 
    geom_bar(stat = "identity") +
    labs(x = "Media Type", y = "Count", title = "Types of Media") +
    facet_wrap(~ type)

```

These graphs convey the number of movies and TV shows for each rating.

```{r}
clean_genres <- str_remove_all(netflix$listed_in, ",.*") 

```

This function eliminates the cross-listing of genres to make for an easier analysis.

```{r}
sum(str_detect(netflix$country, "Singapore|China|India|South Korea|Indonesia|Japan|Taiwan|Philippines") & !is.na(netflix$country))

```

I am curious how much Netflix content comes from Asian production. str_detect lets us know that 1877 of the shows and movies on Netflix are from Asia.

Sources:

This dataset consists of tv shows and movies available on Netflix as of 2019. The dataset is collected from Flixable which is a third-party Netflix search engine.

It was found on the TidyTuesday github; <https://github.com/rfordatascience/tidytuesday/tree/master/data/2021/2021-04-20>
