---
title: "Project 2"
description: |
  Text Analysis
author: Natalie Roston
date: October 3, 2024
format: html
execute:
  warning: false
  message: false
---

```{r}
library(tidytuesdayR)
library(dplyr)
library(stringr)
library(ggplot2)

tuesdata <- tidytuesdayR::tt_load('2021-04-20')
netflix <- tuesdata$netflix

netflix |>
  filter(!is.na(rating), type == "Movie") |>
  group_by(type, rating) |>
  summarize(count = n()) |>
  filter(count > 0) |>
  ggplot(aes(x = rating, y = count)) + 
    geom_bar(stat = "identity") +
    labs(x = "Media Type", y = "Count", title = "Types of Media") +
    theme_minimal()
```

These graphs convey the distrubution of movies for each rating.

```{r}
netflix |>
  mutate(clean_genres = str_remove_all(netflix$listed_in, ",.*"))
```

This function eliminates the cross-listing of genres to make for an easier analysis.

```{r}
r_comedies <- netflix |>
  mutate(clean_genres = str_remove_all(listed_in, ",.*")) |>
  filter(str_detect(clean_genres, "Comedies")) |>
  summarize(percent_r = mean(rating == "R") * 100)
print(r_comedies)
```

10.6% of all comedies are Rated R- I predicted higher!

```{r}
num_the <- length(grep("^The ", netflix$title, value = TRUE))
print(num_the)
```

I was curious how many shows and movies start with "The". I am surprised to find only 982!

```{r}
sum(str_detect(netflix$country, "Singapore|China|India|South Korea|Indonesia|Japan|Taiwan|Philippines") & !is.na(netflix$country))

```

I was wondering how much Netflix content comes from Asian production. str_detect lets us know that 1877 of the shows and movies on Netflix are from Asia.

```{r}
asia <- c("Singapore", "China", "India", "South Korea", "Indonesia", "Japan", "Taiwan", "Philippines")

netflix |>
  filter(country %in% c(asia, "United States", "United Kingdom")) |>
  mutate(country = ifelse(str_detect(country, paste(asia, collapse = "|")), "Asia", country)) |>
  group_by(type, country) |>
  summarize(count = n()) |>
  ggplot(aes(x = type, y = count, fill = country)) +
  geom_bar(stat = "identity") +
  labs(title = "Movies and TV Shows from Asia, UK, and USA ")
      
  
```

This graph conveys the proportion of Asian, UK, and USA production in movies and TV shows. I am surprised at how much Asian production there is, and anticipated more British!

Sources:

This dataset consists of tv shows and movies available on Netflix as of 2019. The dataset is collected from Flixable which is a third-party Netflix search engine.

It was found on the TidyTuesday github; <https://github.com/rfordatascience/tidytuesday/tree/master/data/2021/2021-04-20>
