---
title: "WAI Database SQL Query"
author: Natalie Roston
date: November 26, 2024
format: html
execute:
  warning: false
  message: false
---

To replicate the table in Voss (2020), I will be pulling the measurements, PI info, and subject tables from the scidb.smith.edu database. After making sure I understand the data structure, I will turn the Frequency data to a log scale. I will then select and join the relevant data before preparing it for visualization. The above steps are in SQL, but I will use R's ggplot package to produce the visualization.

```{r, echo=FALSE}
library(RMariaDB)
library(ggplot2)
library(dplyr)
library(mdsr)
library(DBI)
library(tidyverse)


con_wai <- dbConnect(
  MariaDB(), host = "scidb.smith.edu",
  user = "waiuser", password = "smith_waiDB", 
  dbname = "wai"
)
Measurements <- tbl(con_wai, "Measurements")
PI_Info <- tbl(con_wai, "PI_Info")
Subjects <- tbl(con_wai, "Subjects")

# collect(Measurements)
```

```{sql}
#| connection: con_wai

SHOW TABLES;
```

*This shows us the different tables in the database.*

```{sql}
#| connection: con_wai

SELECT *
FROM Measurements
LIMIT 0, 5;
```

*This helped me understand the data structure of Measurements.*

```{sql}
#| connection: con_wai

SELECT * FROM PI_Info
WHERE Identifier = "Abur_2014";
```

*I explored the PI_Info data by looking at a particular study using Identifier; Abur (2014).*

```{sql}
#| connection: con_wai

SELECT * FROM Measurements
WHERE Identifier = "Abur_2014";
```

*I then used the Identifier to explore the corresponding Measurements data.*

```{sql}
#| connection: con_wai

SELECT p.Identifier, Year, AuthorsShortList,
COUNT(DISTINCT SubjectNumber, Ear) AS ears,
CONCAT(AuthorsShortList, " (" , year, ") " , "N=" ,
COUNT(DISTINCT SubjectNumber, Ear), "; ", Instrument) AS Legend

FROM PI_Info AS p

LEFT JOIN Measurements AS m ON m.Identifier = p.Identifier

WHERE p.Identifier IN ("Abur_2014", "Feeney_2017", "Groon_2015" ,"Lewis_2015", 
"Liu_2008", "Rosowski_2012", "Shahnaz_2006", "Shaver_2013" , "Sun_2016", 
"Voss_1994", "Voss_2010", "Werner_2010" ) AND Frequency > 200 AND 
Frequency < 8000

GROUP BY Identifier, Instrument;


```

```{sql}
#| connection: con_wai
#| output.var: "graph"

SELECT p.Identifier, p.Year, p.AuthorsShortList, Frequency,
LOG10(Frequency) AS log_frequency, AVG(Absorbance) AS mean_absorbance,
COUNT(DISTINCT SubjectNumber, Ear) AS ear_u,

CONCAT(AuthorsShortList, " (" , year, ") ", "N=", COUNT(DISTINCT SubjectNumber, Ear), 
"; ", Instrument) AS legend

FROM PI_Info AS p

LEFT JOIN Measurements AS m ON m.Identifier = p.Identifier

WHERE p.Identifier IN ("Abur_2014", "Feeney_2017", "Groon_2015" ,"Lewis_2015", 
"Liu_2008", "Rosowski_2012", "Shahnaz_2006", "Shaver_2013" , "Sun_2016", "Voss_1994", "Voss_2010", "Werner_2010" ) AND Frequency > 200 AND Frequency < 8000

GROUP BY Identifier, Instrument, Frequency;
```


```{r}
graph |>
  ggplot(aes (x= Frequency, y = mean_absorbance,
  color = legend,
  group = legend)) +
  geom_line() +
  scale_x_log10() +
  labs(
    x = "Frequency (Hz)",
    y = "Mean Absorbance",
    title = "Mean Absorbance from Each Publication in WAI Database",
    color = "Legend")

```

*This graph is created by joining Measurements with p.Identifier. There appears to be a strong positive correlation between Frequency and Mean Absorbance overall, but it differs between the studies.*

Question 2:

```{sql}
#| connection: con_wai
SELECT *
FROM Subjects
```

*I wanted to examine the Subjects dataframe to determine which studies had participants of diverse races.*


```{sql}
#| connection: con_wai
#| output.var: "q2_graph"

SELECT Frequency, Race,
    AVG(Absorbance) AS ave_absorbance
FROM Subjects AS s
INNER JOIN Measurements AS m ON m.SubjectNumber = s.SubjectNumber
WHERE m.Identifier = "Rosowski_2012" AND s.Identifier = "Rosowski_2012" 
AND Frequency > 200 AND Frequency < 8000
GROUP BY Race, Frequency;
```


```{r}
q2_graph |>
  ggplot(aes(x = Frequency, y = ave_absorbance, color = Race, group = Race)) +
  geom_line() +
  labs(
    x = "Frequency (Hz)",
    y = "Mean Absorbance",
    title = "Mean Absorbance from Rosowski et al across Race",
    color = "Race"
  ) +
  scale_x_log10()
```

*Interestingly, Rosowski (2012) separated "Chinese" from "Asian" and had very little racial diversity overall. The graphs convey that after a strong positive correlation, average absorbance rapidly dropped after Frequency of 3000.*

For Question 1, it was important that I aggregate data for specific studies with Identifier and that I filter within the frequency range of 80000 \> Frequency \> 200. For Question 2,I wanted to explore the race variable. After determining that Rosowski (2012) enrolled multiple race groups, I joined the relevant variables using the "Rosowksi_2012" Identifier. I then graphed the data, grouping by race.

Article Citations:

Voss SE. Resource Review. Ear Hear. 2019 Nov/Dec;40(6):1481. doi: 10.1097/AUD.0000000000000790. PMID: 31651606; PMCID: PMC7093226.

Rosowski, J. J., Nakajima, H. H., Hamade, M. A., Mahfoud, L., Merchant, G. R., Halpin, C. F., & Merchant, S. N. (2012). Ear-canal reflectance, umbo velocity, and tympanometry in normal-hearing adults. *Ear and Hearing, 33*(1), 19–34. <https://doi.org/10.1097/AUD.0b013e31822ccb76>

Database Source: <https://www.science.smith.edu/wai-database/>
