---
title: "WAI Database SQL Query"
---
To replicate the table in Voss (2020), I will be pulling the measurements, PI info, and subject tables from the scidb.smith.edu database. After making sure I understand the data structure, I will turn the Frequency data to a log scale. I will then select and join the relevant data before preparing it for visualization. The above steps are in SQL, but I will use R's ggplot package to produce the visualization.  

```{r}
library(RMariaDB)
library(ggplot2)
library(dplyr)
library(mdsr)
library(DBI)
library(tidyverse)


con_wai <- dbConnect(
  MariaDB(), host = "scidb.smith.edu",
  user = "waiuser", password = "smith_waiDB", 
  dbname = "wai"
)
Measurements <- tbl(con_wai, "Measurements")
PI_Info <- tbl(con_wai, "PI_Info")
Subjects <- tbl(con_wai, "Subjects")

# collect(Measurements)
```

```{sql}
#| connection: con_wai

SHOW TABLES;
```

```{sql}
#| connection: con_wai

SELECT *
FROM Measurements
LIMIT 0, 5;
```

```{sql}
#| connection: con_wai

SELECT *
FROM PI_Info
LIMIT 0, 50;
```

```{sql}
#| connection: con_wai

SELECT * FROM PI_Info
WHERE Identifier = "Abur_2014";
```

```{sql}
#| connection: con_wai

SELECT * FROM Measurements
WHERE Identifier = "Abur_2014";
```

```{sql}
#| connection: con_wai
#| output.var: "graph"

SELECT Identifier, Frequency, LOG10(Frequency) AS log_frequency, AVG(Absorbance) AS mean_absorbance

FROM Measurements

WHERE Identifier IN ("Abur_2014", "Feeney_2017", "Groon_2015" ,"Lewis_2015", "Liu_2008", "Rosowski_2012", "Shahnaz_2006", "Shaver_2013" , "Sun_2016", "Voss_1994", "Voss_2010", "Werner_2010" ) AND Frequency > 200 AND Frequency < 8000

GROUP BY Identifier, Frequency;
```

```{r}
graph |>
  ggplot(aes (x=Frequency, y = mean_absorbance,
  color = Identifier,
  group = Identifier)) +
  geom_line()+
  scale_x_log10()

```

```{sql}
#| connection: con_wai

SELECT p.Identifier, p.Year, p.AuthorsShortList,
COUNT(DISTINCT SubjectNumber, Ear) AS ear_you

FROM PI_Info AS p

LEFT JOIN Measurements AS m ON m.Identifier = p.Identifier

WHERE p.Identifier IN ("Abur_2014", "Feeney_2017", "Groon_2015" ,"Lewis_2015", "Liu_2008", "Rosowski_2012", "Shahnaz_2006", "Shaver_2013" , "Sun_2016", "Voss_1994", "Voss_2010", "Werner_2010" ) AND Frequency > 200 AND Frequency < 8000

GROUP BY Identifier, Instrument;


```

```{sql}
#| connection: con_wai

SELECT p.Identifier, Year, AuthorsShortList,
COUNT(DISTINCT SubjectNumber, Ear) AS ears,
CONCAT(AuthorsShortList, " (" , year, ") " , "N=" ,
COUNT(DISTINCT SubjectNumber, Ear), "; ", Instrument) AS Legend

FROM PI_Info AS p

LEFT JOIN Measurements AS m ON m.Identifier = p.Identifier

WHERE p.Identifier IN ("Abur_2014", "Feeney_2017", "Groon_2015" ,"Lewis_2015", "Liu_2008", "Rosowski_2012", "Shahnaz_2006", "Shaver_2013" , "Sun_2016", "Voss_1994", "Voss_2010", "Werner_2010" ) AND Frequency > 200 AND Frequency < 8000

GROUP BY Identifier, Instrument;


```

```{sql}
#| connection: con_wai
#| output.var: "graph2"

SELECT p.Identifier, p.Year, p.AuthorsShortList, Frequency,
LOG10(Frequency) AS log_frequency, AVG(Absorbance) AS mean_absorbance,
COUNT(DISTINCT SubjectNumber, Ear) AS ear_u,

CONCAT(AuthorsShortList, " (" , year, ") ", "N=", COUNT(DISTINCT SubjectNumber, Ear), "; ", Instrument) AS legend

FROM PI_Info AS p

LEFT JOIN Measurements AS m ON m.Identifier = p.Identifier

WHERE p.Identifier IN ("Abur_2014", "Feeney_2017", "Groon_2015" ,"Lewis_2015", "Liu_2008", "Rosowski_2012", "Shahnaz_2006", "Shaver_2013" , "Sun_2016", "Voss_1994", "Voss_2010", "Werner_2010" ) AND Frequency > 200 AND Frequency < 8000

GROUP BY Identifier, Instrument, Frequency;
```

```{r}
graph2 |>
  ggplot(aes (x=Frequency, y = mean_absorbance,
  color = legend,
  group = legend)) +
  geom_line()+
  scale_x_log10()

```

Question 2:

```{sql}
#| connection: con_wai
SELECT *
FROM Subjects
```
```{sql}
#| connection: con_wai
#| output.var: "q2_graph"

SELECT Frequency, Race,
    AVG(Absorbance) AS ave_absorbance
FROM Subjects AS s
RIGHT JOIN Measurements AS m ON m.Identifier = s.Identifier 
WHERE m.Identifier = "Rosowski_2012" AND Frequency > 200 AND Frequency < 8000
GROUP BY Race, Frequency
```


```{r}
head(q2_graph)
```
```{sql}
#| connection: con_wai

SELECT race, COUNT(*)
FROM Subjects WHERE Identifier = "Rosowski_2012"
GROUP BY Race
```

```{r}
q2_graph |>
  ggplot(aes (x = Frequency, y = ave_absorbance,
              color = Race,
              group = Race)) +
  geom_line() +
  facet_wrap(~ Race) + 
  scale_x_log10()
```
*Because data was identical across races, I used facet_wrap to isolate the graphs by Race. Interestingly, Rosowski (2012) separated "Chinese" from "Asian" and had very little racial diversity overall. The graphs convey that after a strong positive correlation, average absorbance rapidly dropped after Frequency of 3000. 

For Question 1, it was important that I aggregate data for specific studies with Identifier and that I filter within the frequency range of 80000 > Frequency > 200. For Question 2,I wanted to explore the race variable. After determiniing that Rosowski (2012) enrolled multiple race groups, I joined the relevant variables using the "Rosowksi_2021" Identifier. I then graphed the data, grouping by race. 

