---
title: "Refugees in Germany Over Time"
description: |
  Analysis of refugee data from TidyTuesday
author: Natalie Roston
date: September 18, 2024
format: html
execute:
  warning: false
  message: false
---
This analysis examines refugee and asylum seeker trends in Germany using 
TidyTuesday data. The visualizations track changes over time, highlighting the 
distinct patterns for each group. These findings provide insight into migration 
dynamics and Germany’s role in global displacement trends.

```{r}
library(dplyr)

tuesdata <- tidytuesdayR::tt_load('2023-08-22')
population <- tuesdata$population
desired_variables <- population |>
     select(year, coa_name, refugees, asylum_seekers)
germany_data <- desired_variables |>
     filter(coa_name == "Germany")
summary_germany <- germany_data |>
     group_by(year) |>
     summarize(
         total_refugees = sum(refugees, na.rm = TRUE),
         total_asylum_seekers = sum(asylum_seekers, na.rm = TRUE),
         .groups = 'drop'
    )

library(ggplot2)
ggplot(summary_germany, aes(x = year)) +
     geom_line(aes(y = total_refugees, color = "Refugees"), size = 1) +
     geom_line(aes(y = total_asylum_seekers, color = "Asylum Seekers"), size = 1) +
     labs(
         title = "Total Refugees and Asylum Seekers in Germany by Year",
         x = "Year",
         y = "Count",
         color = "Legend"
     ) +
     theme_minimal()
```

Sources:

Data Science Learning Community (2024). Tidy Tuesday: A weekly social data project. [https://tidytues.day](https://tidytues.day/)

<https://github.com/rfordatascience/tidytuesday/tree/master/data/2023/2023-08-22>

-   Data from UNHCR’s annual statistical activities dating back to 1951.

-   Data from the United Nations Relief and Works Agency for Palestine Refugees 
in the Near East (UNRWA), specifically for registered Palestine refugees under 
UNRWA’s mandate.

-   Data from the Internal Displacement Monitoring Centre (IDMC) on people 
displaced within their country due to conflict or violence

````{=html}
<!--
```{r}
library(dplyr)
library(tidytuesdayR)
tuesdata <- tidytuesdayR::tt_load('2023-08-22')
population <- tuesdata$population


desired_variables <- population %>%
     select(year, coa_name, refugees, asylum_seekers)
```

-->
````
